Temporal ordering and registration of images in studies of developmental dynamics

### 2015-01-25 image_analysis_paper_development2/manuscript.tex
added development paper text

\begin{figure}[t]
\includegraphics{fig1}
\caption{Caricature illustrating the tasks of image registration and temporal ordering. (A) Images of ``samples'', each in a different orientation and a different stage of development. (B) Registered and ordered samples. For this caricature, the registration and ordering is straightforward because the data set is small, the landmarks are visually apparent, and the developmental changes are easy to recognize.}
%\label{fig:fish}
\customlabel{fig:fish}{1}
\customlabel{subfig:fish_unordered}{\ref{fig:fish}A}
\customlabel{subfig:fish_ordered}{\ref{fig:fish}B}
\end{figure}

@@

\begin{figure}[t]
\includegraphics{fig2}
\caption{Schematic illustrating angular synchronization and diffusion maps. (A) Set of vectors, each in a different orientation. The pairwise alignment angles are indicated. (B) The vectors from A, each rotated about their midpoint so that the set is globally aligned. Note that the chosen rotation angles are consistent with the pairwise alignments in A: the differences between a pair of angles in B is the same as the pairwise angle in A. (C) Data points (in black) which lie on a one-dimensional nonlinear curve in two dimensions. Each pair of points is connected by an edge, and the edge weight is related to the Euclidean distance between the points through the diffusion kernel (see \SI), so that close data points are connected by darker (``stronger'') edges. (D) The data in C, colored by the first (non-trivial) eigenvector from the diffusion map computational procedure. The color intensity is monotonic with the perceived curve arclength, thus parameterizing the curve.}
%\label{fig:schematics}
\customlabel{fig:schematics}{2}
\customlabel{subfig:synch1}{\ref{fig:schematics}A}
\customlabel{subfig:synch2}{\ref{fig:schematics}B}
\customlabel{subfig:dmaps1}{\ref{fig:schematics}C}
\customlabel{subfig:dmaps2}{\ref{fig:schematics}D}
\end{figure}

@@

\begin{figure*}[t]
\includegraphics{fig3}
\caption{Method validation using live imaging of \textit{Drosophila} embryos. (A) Selected images from a live imaging study of a \textit{Drosophilla} embryo during gastrulation. Each frame is in an arbitrary rotational orientation, and the order of the frames has been shuffled. (B) Images from A after registration and ordering using vector diffusion maps. The dorsal side of each embryo now appears at the top of each image, and the ventral side appears at the bottom. (C) The correlation between the recovered rotation angle (using vector diffusion maps) and the true rotation angle. (D) The correlation between the recovered rank (using vector diffusion maps) and the true rank. (E) The average error in the recovered angle and the rank correlation coefficient for $5$ different live imaging studies. (F) The median rank correlation coefficient (over $200$ samples) as a function of the number of images in the data set. }
\customlabel{fig:movie}{3}
\customlabel{subfig:movie_unregistered_unordered}{\ref{fig:movie}A}
\customlabel{subfig:movie_registered_ordered}{\ref{fig:movie}{B}}
\customlabel{subfig:movie_angle_corr}{\ref{fig:movie}{C}}
\customlabel{subfig:movie_rank_corr}{\ref{fig:movie}{D}}
\customlabel{subfig:movie_error_table}{\ref{fig:movie}{E}}
\customlabel{subfig:movie_bootstrap}{\ref{fig:movie}{F}}
\end{figure*}

@@

@@

\subsection*{Data sets with interembryo variability}

@@

\begin{figure*}[t]
\includegraphics{fig4}
\caption{Analyzing images of fixed \textit{Drosophila} embryos. {(A)} Images of \textit{ Drosophila} embryos, stained for nuclei (gray), Twi (green), and dpERK (red). Each image is of a different embryo arrested at a different developmental time and in a different rotational orientation. {(B)} Data from {\it A}, registered and ordered using vector diffusion maps. The expert rank for each image is indicated. {(C)} A representative ``developmental trajectory'' obtained from local averaging of the entire set of registered and ordered images (see \SI). {(D)} Correlation between the image ranks calculated from the vector diffusion maps algorithm and the ranks obtained from ordering by an expert. The rank correlation coefficient is $0.97$. }
\customlabel{fig:data2}{4}
\customlabel{subfig:fixed_images_unregistered_unordered}{\ref{fig:data2}{A}}
\customlabel{subfig:fixed_images_registered_ordered}{\ref{fig:data2}{B}}
\customlabel{subfig:fixed_images_average_trajectory}{\ref{fig:data2}{C}}
\customlabel{subfig:fixed_images_rank_corr}{\ref{fig:data2}{D}}
\end{figure*}

>>>>

%\begin{figure}[t]
%\includegraphics{fig1}
%\caption{Caricature illustrating the tasks of image registration and temporal ordering. (A) Images of ``samples'', each in a different orientation and a different stage of development. (B) Registered and ordered samples. For this caricature, the registration and ordering is straightforward because the data set is small, the landmarks are visually apparent, and the developmental changes are easy to recognize.}
%%\label{fig:fish}
%\customlabel{fig:fish}{1}
%\customlabel{subfig:fish_unordered}{\ref{fig:fish}A}
%\customlabel{subfig:fish_ordered}{\ref{fig:fish}B}
%\end{figure}

@@

%\begin{figure}[t]
%\includegraphics{fig2}
%\caption{Schematic illustrating angular synchronization and diffusion maps. (A) Set of vectors, each in a different orientation. The pairwise alignment angles are indicated. (B) The vectors from A, each rotated about their midpoint so that the set is globally aligned. Note that the chosen rotation angles are consistent with the pairwise alignments in A: the differences between a pair of angles in B is the same as the pairwise angle in A. (C) Data points (in black) which lie on a one-dimensional nonlinear curve in two dimensions. Each pair of points is connected by an edge, and the edge weight is related to the Euclidean distance between the points through the diffusion kernel (see \SI), so that close data points are connected by darker (``stronger'') edges. (D) The data in C, colored by the first (non-trivial) eigenvector from the diffusion map computational procedure. The color intensity is monotonic with the perceived curve arclength, thus parameterizing the curve.}
%%\label{fig:schematics}
%\customlabel{fig:schematics}{2}
%\customlabel{subfig:synch1}{\ref{fig:schematics}A}
%\customlabel{subfig:synch2}{\ref{fig:schematics}B}
%\customlabel{subfig:dmaps1}{\ref{fig:schematics}C}
%\customlabel{subfig:dmaps2}{\ref{fig:schematics}D}
%\end{figure}

@@

%\begin{figure*}[t]
%\includegraphics{fig3}
%\caption{Method validation using live imaging of \textit{Drosophila} embryos. (A) Selected images from a live imaging study of a \textit{Drosophilla} embryo during gastrulation. Each frame is in an arbitrary rotational orientation, and the order of the frames has been shuffled. (B) Images from A after registration and ordering using vector diffusion maps. The dorsal side of each embryo now appears at the top of each image, and the ventral side appears at the bottom. (C) The correlation between the recovered rotation angle (using vector diffusion maps) and the true rotation angle. (D) The correlation between the recovered rank (using vector diffusion maps) and the true rank. (E) The average error in the recovered angle and the rank correlation coefficient for $5$ different live imaging studies. (F) The median rank correlation coefficient (over $200$ samples) as a function of the number of images in the data set. }
%\customlabel{fig:movie}{3}
%\customlabel{subfig:movie_unregistered_unordered}{\ref{fig:movie}A}
%\customlabel{subfig:movie_registered_ordered}{\ref{fig:movie}{B}}
%\customlabel{subfig:movie_angle_corr}{\ref{fig:movie}{C}}
%\customlabel{subfig:movie_rank_corr}{\ref{fig:movie}{D}}
%\customlabel{subfig:movie_error_table}{\ref{fig:movie}{E}}
%\customlabel{subfig:movie_bootstrap}{\ref{fig:movie}{F}}
%\end{figure*}

@@

\subsubsection*{{\em Drosophila} gastrulation}

\subsubsection*{Zebrafish gastrulation}

@@

\subsection*{Data sets with intersample variability}

\subsubsection*{Fixed {\em Drosophila} images}

\subsubsection*{Wing disc data}

@@

%\begin{figure*}[t]
%\includegraphics{fig4}
%\caption{Analyzing images of fixed \textit{Drosophila} embryos. {(A)} Images of \textit{ Drosophila} embryos, stained for nuclei (gray), Twi (green), and dpERK (red). Each image is of a different embryo arrested at a different developmental time and in a different rotational orientation. {(B)} Data from {\it A}, registered and ordered using vector diffusion maps. The expert rank for each image is indicated. {(C)} A representative ``developmental trajectory'' obtained from local averaging of the entire set of registered and ordered images (see \SI). {(D)} Correlation between the image ranks calculated from the vector diffusion maps algorithm and the ranks obtained from ordering by an expert. The rank correlation coefficient is $0.97$. }
%\customlabel{fig:data2}{4}
%\customlabel{subfig:fixed_images_unregistered_unordered}{\ref{fig:data2}{A}}
%\customlabel{subfig:fixed_images_registered_ordered}{\ref{fig:data2}{B}}
%\customlabel{subfig:fixed_images_average_trajectory}{\ref{fig:data2}{C}}
%\customlabel{subfig:fixed_images_rank_corr}{\ref{fig:data2}{D}}
%\end{figure*}
